---
title: "What add_lmed_v20260902() does"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{What add_lmed_v20260902() does}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(data.table)
library(mht)

# Sort category codes as A1, A2, ..., A7, B1, B2, ..., B12, C1, ...
cat_order <- function(x) {
  x <- unique(x)
  return(x[order(substr(x, 1, 1), as.integer(substring(x, 2)))])
}

# One row per run of equal values in one column, per person.
episodes <- function(skeleton, col) {
  out <- skeleton[,
    .(from = isoyearweek[1], to = isoyearweek[.N], weeks = .N, level = get(col)[1]),
    by = .(id, run = rleid(get(col)))
  ]
  out[, run := NULL]
  return(out[])
}

# A one-person skeleton of consecutive ISO weeks, starting Monday 2020-01-06.
monday <- as.Date("2020-01-06")
toy_skeleton <- function(n_weeks = 40L) {
  return(data.table(
    id = 1L,
    isoyearweek = cstime::date_to_isoyearweek_c(monday + 7L * (seq_len(n_weeks) - 1L))
  ))
}

# One prescription for person 1, dispensed on the Monday of `start_week`,
# covering `n_weeks` whole ISO weeks.
rx <- function(produkt, start_week, n_weeks, lnmn = NA_character_) {
  return(data.table(
    lopnr = 1L,
    produkt = produkt,
    edatum = monday + 7L * (start_week - 1L),
    fddd = 7 * n_weeks,
    lnmn = lnmn
  ))
}
```

`add_lmed_v20260902()` takes a table of dispensed prescriptions and a
person-week skeleton. It writes menopausal hormone therapy (MHT) exposure
columns and two person-level exclusion columns onto the skeleton. Exposure
here means dispensed supply, assumed used. The function reads the dispensed
quantity and, for two strength-keyed rules, the strength. It reads no
adherence and no indication. This vignette states what it does, in the order
it does it:

1. It looks each product name up in a shipped product table. The table gives
   the product's MHT category, and whether the study removes its holders.
2. It turns each prescription into a closed interval of ISO weeks.
3. It marks each person-week with the categories in force that week.
4. It resolves the categories into one treatment level per week, under each of
   three approaches.
5. It derives the `rd_approach*` exposure variables from those levels.
6. It flags every week of a person whose supplied prescriptions include a
   product the table marks for removal.

The product table is `product_table_20260902.xlsx`. The duration rules and the
approach rules come from the codebook `dataDictionary20260828.xlsx`. Both
files ship with the package, and the tables below read them when the vignette
builds, so they cannot disagree with either file. Steps 2 to 5 are the
2026-08-28 layer, reached through that layer's own dated helpers. Section 1
lists the products the two entry points classify differently, and section 3
verifies that they agree on the fixture. The function's arguments, the
columns it owns and the conditions it rejects are in `?add_lmed_v20260902`.
This vignette does not repeat them.

```{r files}
product_table <- system.file("2023-mht", "product_table_20260902.xlsx", package = "mht")
codebook <- system.file("2023-mht", "dataDictionary20260828.xlsx", package = "mht")
sheets <- setdiff(readxl::excel_sheets(product_table), "README")
sheets
ptab <- rbindlist(
  lapply(sheets, function(s) {
    setDT(suppressMessages(readxl::read_excel(product_table, sheet = s, col_types = "text")))
  }),
  use.names = TRUE,
  fill = TRUE
)
ptab[, .N, keyby = .(sheet = fifelse(atc %like% "^G0[23]", "G", "other"))]
groups <- setDT(readxl::read_excel(codebook, sheet = "MHT_groups", col_types = "text"))
groups <- groups[!is.na(Preparatnamn)]
```

## 1. Products and categories

### How a product name is looked up

The function reduces the register's product name to its ASCII letters, then
lowercases it. `Estradot 25`, `Estradot®` and `Estradot` all become
`estradot`. It then looks that key up in the product table, on an exact match.
Extra digits, spaces and symbols disappear in the reduction, so they still
match. Extra letters do not: `estradotmatrix` would match no key. The table
holds one row per raw register name, so several rows can share one key. Rows that share a key MUST
agree on the category, on the exclusion flag and on its reason. The function
errors where they do not.

A delivered name with no row is an error. The function checks the whole
prescription table before it restricts to the people in the skeleton. A
product held only by people outside the cohort therefore still needs a row. A
product nobody has ruled on would otherwise read as no MHT, and its holder
would enter as an unexposed control. The table was built from the two
register deliveries the studies hold, so a later delivery can bring a name it
lacks. The function then stops until the table gains the row.

`notmht` in the table is a recorded decision that the product is not MHT. A
prescription of such a product contributes nothing, and the function reports
nothing about it.

The reader validates four things. A category MUST be one the person-week
grid materialises, one the 2026-08-28 classifier returns, or `notmht`. The
exclusion cell MUST read `TRUE` or `FALSE`. The key MUST equal the raw name
reduced, so a key is never trusted. Rows sharing a key MUST agree, as above.
No other cell is checked.

### The categories

The table names 33 categories, each with at least one product. A category is
one letter for the drug group and one number for a subgroup: the route, the
progestogen or the regimen. The table below lists every product that reaches
each category. The `meaning` column is the product table's own English
translation of the codebook's group and route cells, shown without the
table's parenthetical notes. The route is the codebook's. The lookup reads
the product name only, and never checks the route dispensed.

```{r legend}
legend <- ptab[classification != "notmht",
  .(
    meaning = paste(unique(sub(" \\(.*\\)$", "", classification_meaning)), collapse = " | "),
    products = paste(unique(produkt_raw), collapse = ", ")
  ),
  keyby = .(category = classification)
]
legend <- legend[order(substr(category, 1, 1), as.integer(substring(category, 2)))]
knitr::kable(legend)
```

The letter groups the drug. `A` is oestrogen alone. `B` is oestrogen with a
progestogen in one product. `C` is a progestogen alone. `D`, `E` and `I` are
contraceptive progestogens. `F` is tibolone, `G` is a tissue-selective
oestrogen complex and `H` is an androgen. The codebook counts `A3`, `A4` and
`A5` as local treatment, although `A5` is a peroral oestriol. No approach rule
treats one of the three as MHT exposure.

### Products that are not MHT

The table's `G` sheet covers the ATC groups G02 and G03. These are its
`notmht` rows:

```{r notmht}
ptab[classification == "notmht" & atc %like% "^G0[23]", .(atc, produkt_raw)]
```

The `other` sheet covers every other ATC group the deliveries hold. Every row
of it is `notmht`:

```{r other}
other <- ptab[!atc %like% "^G0[23]"]
nrow(other)
other[, .N, keyby = .(classification, exclude_entire_person)]
```

### What changed against the 2026-08-28 classifier

The 2026-08-28 entry point classifies with a prefix ladder written into the
code. The chunk below runs that ladder over every raw name of the `G` sheet
and lists the names where the table answers differently. A ladder answer of
`none` and a table answer of `notmht` are the same outcome, and are not
listed.

```{r changed}
compare <- ptab[atc %like% "^G0[23]", .(produkt = produkt_raw, table = classification)]
mht:::lmed_categorize_product_names_v20260828(compare)
compare[, ladder := fifelse(is.na(product_category), "notmht", product_category)]
compare[table != ladder, .(produkt, ladder, table)]
```

Each of these products reached no category under the ladder and reaches one
under the table. Endovelle is now a `C3` product. In a week it shares with a
systemic oestrogen it lights `estrogen_progesterone_synthetic`, and it keeps
`estrogen_only` from lighting.

### Categories that never change exposure

Two categories can be classified and still move no exposure variable, `G1`
and `H1`. A third, `D4`, is a column that no product reaches. `A3`, `A4` and
`A5` are read only by the inert reference-level rules of section 4, so they
move nothing either.

```{r unread}
pg <- setDT(suppressMessages(
  readxl::read_excel(codebook, sheet = "post_grouping", col_types = "text")
))
pg <- pg[!is.na(approach)]
ex_cols <- grep("^doesnotinclude", names(pg), value = TRUE)
read <- cat_order(na.omit(unlist(pg[, c("includes1", "includes2", ex_cols), with = FALSE])))
read
```

- `G1` is Duavive alone, and Duavive counts as no MHT. The table gives it
  `G1`, and the person-week grid has no `G1` column, so no rule can read it.
- `H1` is the androgens. The grid carries an `H1` column, and no approach rule
  names it.
- `D4` appears nowhere in the codebook. The grid carries a `D4` column that is
  `FALSE` in every week, and no rule names it.

### Products whose holders the study removes

The table flags a product with `exclude_entire_person`. The flag is separate
from the category: a flagged product keeps its category, so a reversed
exclusion loses nothing. Section 6 states what the function does with the
flag.

```{r excluded}
ptab[exclude_entire_person == "TRUE", .(produkt_raw, classification, exclusion_reason)]
```

## 2. From a prescription to an interval

The layer reads five columns of the prescription table: the person identifier,
`produkt`, `edatum`, `fddd`, and `lnmn` where a strength-keyed rule needs it.
It reads `fddd` as a number of days, rounded to a whole day with `round()`,
unless a codebook rule says otherwise.

Four steps run, in this order:

1. Look the product up, as in section 1.
2. Overwrite `fddd` with the codebook's fixed duration, for the products that
   carry one.
3. Drop every prescription with no category, no dispensing date, or no
   positive duration.
4. Read the strength from `lnmn`, and apply the whole-months rules.

Step 2 runs before step 3 because an implant carries no useful dispensed
quantity, and the fixed duration supplies one. Step 4 runs after step 3
because a prescription with no duration has no interval, whatever its
strength. Reading the strength first would turn a missing `lnmn` on such a row
into an error for nothing.

The duration rules come from the codebook's `MHT_groups` sheet, and they are
matched by prefix, not by exact key. A rule applies where its codebook name is
a prefix of the reduced product name, and the longest such name wins.
`Utrogestan 100mg Capsules` takes the `Utrogestan` rule. An alias in the
duration layer gives `Utrogest` and `Gepretix` the `Utrogestan` rules. Their
categories come from the table, where all three are `C1`. The category lookup
and the duration rule are chosen independently, and nothing checks that the
rule's codebook subgroup agrees with the table's category.

The interval is closed. A prescription dispensed on `edatum` for `d` days
covers `edatum` to `edatum + d - 1`. Both ends convert to an ISO week with
`cstime::date_to_isoyearweek_c()`. Each prescription gets its own interval.
Overlapping intervals of one category are unioned, and supply left over at a
refill is not carried forward.

### Fixed durations

```{r fixed}
knitr::kable(groups[!is.na(FDDD), .(Preparatnamn, Subgrupp, FDDD)])
```

1680 days is five years of twelve 28-day months. 1008 days is three years of
the same months. Jaydess is the one product whose fixed duration differs from
the rest of its category. The function reads no removal or replacement date,
so a device's interval runs its full length or ends with follow-up.

### Whole months for sparse progestogens

A sequential progestogen is taken on some days of each month. Its `fddd`
counts the days of supply, and a sparse supply covers more calendar time than
it has days. For each such product the codebook gives `minimum_monthly_dose`,
the days of supply that make one treatment month, and `minimum_months`, the
months a prescription must reach.

```{r months}
knitr::kable(groups[
  !is.na(minimum_monthly_dose),
  .(Preparatnamn, Subgrupp, minimum_monthly_dose, minimum_months, strength_mg_min, strength_mg_max)
])
```

The rule rewrites `fddd` in three steps:

1. `months = floor(fddd / minimum_monthly_dose)`.
2. If `months < minimum_months`, the duration is zero and the prescription is
   dropped.
3. Otherwise `fddd = months * 28`.

A month is 28 days because the codebook says so and the grid counts whole
weeks. The result rounds down, never up.

Two Utrogestan rules are keyed on strength. `strength_mg_min` is inclusive
and `strength_mg_max` is exclusive, so the pair `NA, 150` and `150, NA` covers
every strength once. The strength comes from `lnmn`, the register's full
product name, which carries the dosage form and the strength. The strength is
the last number in `lnmn` followed by `mg`, with a comma read as a decimal
point. A number followed by `mg/` is a concentration and does not count. As
soon as a strength-keyed product survives step 3, `lnmn` is required. A
surviving row of it with no readable strength is an error.

The chunk below drives the duration layer with one row per case. The layer is
an internal function, so the chunk calls it directly.

```{r durations}
cases <- data.table(
  lopnr = 1L,
  produkt = c("Divigel", "Cerazette", "Provera", "Provera", "Utrogestan", "Utrogestan", "Utrogest", "Mirena", "Jaydess"),
  lnmn = c(NA, NA, NA, NA, "Utrogestan, kapsel, mjuk 100 mg", "Utrogestan, kapsel, mjuk 200 mg", "Utrogest, kapsel, mjuk 100 mg", NA, NA),
  edatum = as.Date("2020-01-06"),
  fddd = c(90, 28, 28, 8, 30, 30, 30, 1, 1)
)
intervals <- mht:::lmed_durations_v20260902(cases, verbose = FALSE)
intervals[, .(produkt, lnmn, fddd, duration_days, start_isoyearweek, stop_isoyearweek)]
```

Divigel has no rule and keeps 90 days. Cerazette and Provera at 28 days of
supply reach seven months, which is 196 days. Provera at 8 days of supply
reaches two months. The rule needs three, so the prescription is dropped and
the warning reports it. Utrogestan 100 mg needs 28 days of supply per month
and gets one month. Utrogestan 200 mg needs 12 and gets two. Utrogest 100 mg
takes the Utrogestan rule. Mirena and Jaydess take their fixed durations,
whatever `fddd` says.

### Dropped prescriptions

A prescription of a `notmht` product reaches no category and is dropped
silently. Every other drop is counted in one aggregate warning, under three
causes:

- no dispensing date
- no duration
- a duration that is no positive number of days, which includes a whole-months
  result of zero

Outside the six fixed-duration products, a missing `fddd` is never imputed.

## 3. Person-weeks

The skeleton MUST hold each person's ISO weeks once each, with no gap between
the first and the last. Gap bridging and the three-year rule both count rows,
and a row is one elapsed week only under that condition. The function errors
otherwise, before it writes anything.

Each of the 33 grid categories gets one logical column. `G1` is discarded
before this step, and `D4` is a column that no product reaches. A week is
`TRUE` for a category where any interval of that category includes at least
one day of the week. One day of supply on a Sunday marks that week. The same
day plus the Monday marks two weeks. An episode already running at a person's
first week is clipped: it reads as starting in that week. The function counts
observed weeks and never reaches behind the first one.

The synthetic fixtures show the result. `fake_lmed_2026` holds one Utrogestan
user, so `lnmn` is set for those rows first. It also holds one name,
`Paracetamol`, that exists in no real delivery and so has no row in the
product table. That row is removed, or the function stops.

```{r fixture}
skeleton <- copy(fake_skeleton_mht)
lmed <- copy(fake_lmed_2026)
lmed[, lnmn := NA_character_]
lmed[produkt == "Utrogestan", lnmn := "Utrogestan, kapsel, mjuk 100 mg"]
lmed <- lmed[produkt != "Paracetamol"]
add_lmed_v20260902(skeleton, lmed, verbose = FALSE)
categories <- grep("^[A-I][0-9]+$", names(skeleton), value = TRUE)
categories
skeleton[, lapply(.SD, sum), .SDcols = categories][, .SD, .SDcols = function(x) x > 0]
```

The grid materialises 33 categories. Two of them, `D4` and `H1`, are read by
no approach rule:

```{r grid-unread}
setdiff(categories, read)
```

On this input the two entry points agree on every column they share. The
fixture holds none of the products of section 1 that the two classify
differently. The 2026-09-02 function adds the two exclusion columns of
section 6 and nothing else.

```{r agree}
older <- copy(fake_skeleton_mht)
suppressWarnings(add_lmed_v20260828(older, lmed, verbose = FALSE))
shared <- intersect(names(skeleton), names(older))
identical(skeleton[, ..shared], older[, ..shared])
setdiff(names(skeleton), names(older))
```

### The skeleton must be weekly

`swereg::create_skeleton()` writes one annual row per person per year before
the weekly rows. An annual row holds `is_isoyear = TRUE` and an `isoyearweek`
that is no ISO week:

| id | isoyearweek | is_isoyear |
|---|---|---|
| 1 | `1946-**` | `TRUE` |
| 1 | `2015-02` | `FALSE` |

The function stops on the annual row, before it writes anything:

```
skeleton$isoyearweek holds a value that is no ISO week: 1946-**
```

Pass the weekly rows only, and fill the annual rows in the caller:

```r
weekly <- skeleton[is_isoyear == FALSE]
add_lmed_v20260902(weekly, lmed)
```

The subset is a new `data.table`, so the function writes its columns there.

## 4. From categories to a treatment level

The `post_grouping` sheet defines three approaches. Each approach is a set of
rules. A rule names one treatment variable, the categories a week must carry,
and the categories it must not carry. The rule lights its variable in every
week that carries all of the first set and none of the second. Several rules
can name one variable, and each lights it on its own.

The tables below compress the rules that name at least one category. `alone`
lists the categories that light the variable on their own. `pairs` lists the
combinations that light it, written as `oestrogens + partners`. Every
oestrogen listed pairs with every partner listed. The chunk asserts that the
pairs are exactly that cross, with no pair missing and none repeated.
`excludes` lists the categories a week must not carry. Every rule of one
variable carries the same `excludes`, and the chunk asserts that too.

```{r rule-table}
pg[, excludes := apply(.SD, 1, function(r) paste(cat_order(r[!is.na(r)]), collapse = " ")),
  .SDcols = ex_cols]
rule_table <- function(a) {
  x <- pg[approach == a & !is.na(includes1)]
  out <- x[, {
    alone <- includes1[is.na(includes2)]
    p1 <- includes1[!is.na(includes2)]
    p2 <- includes2[!is.na(includes2)]
    seen <- paste(p1, p2)
    cross <- as.vector(outer(unique(p1), unique(p2), paste))
    stopifnot(anyDuplicated(seen) == 0L, setequal(seen, cross))
    stopifnot(length(unique(excludes)) == 1L)
    .(
      rules = .N,
      alone = paste(cat_order(alone), collapse = " "),
      pairs = if (length(p1) == 0L) "" else
        paste(paste(cat_order(p1), collapse = "/"), "+", paste(cat_order(p2), collapse = "/")),
      excludes = paste(unique(excludes[nzchar(excludes)]), collapse = " | ")
    )
  }, keyby = variable]
  return(out[order(variable == "local_or_none_mht", variable)])
}
```

### Approach 1: any systemic MHT

```{r approach-1}
knitr::kable(rule_table("1"))
```

Every `A` group systemic oestrogen and every `B` product is one level,
`systemic_mht`. Tibolone is its own level. No rule of approach 1 reads a
progestogen or a contraceptive. A woman on oestrogen with a separate
progestogen is `systemic_mht`. A woman on a progestogen alone is not exposed.

### Approach 2: route of the oestrogen

```{r approach-2}
knitr::kable(rule_table("2"))
```

Approach 2 splits systemic oestrogen by route. Injected oestrogen (`A7`) sits
with the transdermal route. The combined products follow the route of their
oestrogen: the two patches, `B1` and `B7`, are transdermal, and the ten
tablets are peroral. Tibolone is again its own level.

### Approach 3: the progestogen

```{r approach-3}
knitr::kable(rule_table("3"))
```

Approach 3 asks what accompanies the oestrogen. `estrogen_only` is a systemic
oestrogen with no combined product, no progestogen and no contraceptive in the
same week. The two `estrogen_progesterone` levels split the progestogen. The
codebook's `bioidentical` level holds vaginal progesterone (`C1`) and
dydrogesterone (`C5`, and the combined products `B4` and `B11`). Its
`synthetic` level holds every other progestogen, including injected
progesterone (`C2`), whether in a combined product or dispensed separately.
The two names are the codebook's labels, and this vignette does not judge
them. Tibolone is its own level. A progestogen or a contraceptive alone lights
nothing.

Approach 3 also declares `local_or_none_mht` with an empty rule. That rule
lights nothing.

### The reference level

`local_or_none_mht` is the value a week takes when no treatment variable is
lit. The rules that name it, in every approach, are inert. The resolver never
reads the column they light, and it deletes that column with the rest of its
working columns. The name says what such a week holds: a local oestrogen, a
progestogen alone, a contraceptive, an androgen, Duavive, or nothing.

### Bridging a gap

Before the resolver compares treatment variables, it bridges short gaps in
each of them separately. For one variable, a run of up to four weeks in which
it is unlit, between two weeks in which it is lit, becomes lit. A leading or a
trailing run is not a gap, so no woman starts early and none stops late. The
category columns are not bridged: bridging happens inside the approach step.

```{r bridge}
skel <- toy_skeleton()
add_lmed_v20260902(skel, rbind(rx("Divigel", 1, 10), rx("Divigel", 14, 7)), verbose = FALSE)
episodes(skel, "approach1")
skel <- toy_skeleton()
add_lmed_v20260902(skel, rbind(rx("Divigel", 1, 10), rx("Divigel", 16, 5)), verbose = FALSE)
episodes(skel, "approach1")
```

A three-week gap is bridged. A five-week gap is not, and the woman holds two
episodes.

### One level per week: the most recent start wins

A week can light two treatment variables. Concomitance is judged at ISO-week
resolution, so two products dispensed on different days of one week count as
concomitant. The resolver gives each variable a run length, which counts the
weeks since it last turned on, and reports the variable with the shortest run.
The treatment the woman started most recently is the one reported.

The woman below takes Progynon (`A2`) for 30 weeks and adds Divigel (`A1`)
from week 11 for 20 weeks.

```{r switch}
skel <- toy_skeleton()
add_lmed_v20260902(skel, rbind(rx("Progynon", 1, 30), rx("Divigel", 11, 20)), verbose = FALSE)
episodes(skel, "approach1")
episodes(skel, "approach2")
episodes(skel, "approach3")
```

Approach 1 sees one level throughout, because `A1` and `A2` are both
`systemic_mht`. Approach 2 reports `peroral_estrogen` for ten weeks and then
`transdermal_estrogen`, the more recent start, for twenty. Approach 3 sees
`estrogen_only` throughout.

An exclusion works the same way. The woman below takes Divigel for 20 weeks
and Utrogestan 100 mg from week 9 for 8 weeks.

```{r partner}
skel <- toy_skeleton()
add_lmed_v20260902(
  skel,
  rbind(rx("Divigel", 1, 20), rx("Utrogestan", 9, 8, "Utrogestan, kapsel, mjuk 100 mg")),
  verbose = FALSE
)
episodes(skel, "approach3")
```

While `C1` is in force, `estrogen_only` is not lit, because its rule excludes
`C1`, and `estrogen_progesterone_bioidentical` is lit by the pair `A1 + C1`.
When the progesterone ends, `estrogen_only` lights again. Its eight-week gap
is over four weeks, so it is not bridged, and the run restarts.

### Two starts in one week: a clash

Two treatment variables that start in the same week have equal run lengths,
and neither wins. The week is marked `clashingprescriptions`. The mark carries
forward to the last treated week of that episode, and no further. The first
untreated week after it is an ordinary stop.

```{r clash}
skel <- toy_skeleton()
add_lmed_v20260902(skel, rbind(rx("Progynon", 1, 30), rx("Divigel", 1, 20)), verbose = FALSE)
episodes(skel, "approach2")
```

Neither end of that rule is arbitrary. A clash carried to the end of
follow-up would keep a woman who stopped and restarted out of the comparison
for good. A clash ended with the overlap would read her as a new user in the
week the overlap ends, and she was treated throughout.

## 5. The exposure variables

The exposure layer turns each `approach` column into two exposure columns,
`rd_approach<N>_single` and `rd_approach<N>_multiple`. Each holds the
approach's levels, `clashingprescriptions` among them, plus two more:
`previous` and `exclude`. Five steps run per approach and per variant:

1. Mark the move from an active level to `local_or_none_mht` as `previous`.
   Every later untreated week is also `previous`.
2. Find the first week that moves from `previous` back to an active level.
   That is the re-initiation.
3. In the `single` variant only, set every week from the re-initiation onward
   to `exclude`.
4. Rewrite `previous` to `exclude` where the treated episode before it ran
   under 156 weeks, which is three years of 52.
5. Write the result.

`clashingprescriptions` counts as an active level in every step. A woman who
leaves a clash for no treatment is a former user like any other.

The `single` variant allows one treated episode in the observed follow-up.
Everything from a second start onward leaves the analysis. A start before the
first observed week is not seen, so the first observed episode counts as the
first. The `multiple` variant lets her re-enter her new level, and excludes
person-time only through step 4.

### Former user or excluded

Two women in the fixture stop and never restart. Person 1 was treated for 193
weeks, person 2 for 73.

```{r former}
episodes(skeleton[id %in% c(1, 2)], "rd_approach3_single")
```

Person 1 becomes `previous`, a former user. Person 2 becomes `exclude`,
because 73 weeks is under 156. Every week before a woman's first treatment
stays `local_or_none_mht`.

### Re-initiation, single against multiple

Person 6 was treated for 193 weeks, stopped for 21, and restarted for 50.

```{r reinitiation}
episodes(skeleton[id == 6], "rd_approach3_single")
episodes(skeleton[id == 6], "rd_approach3_multiple")
```

In the `single` variant, the 21 untreated weeks are `previous`, because the
first episode reached 156 weeks, and every week from the restart is
`exclude`. In the `multiple` variant she re-enters `estrogen_only` for 50
weeks. Her final untreated weeks are `exclude`, because that second episode
ran under 156 weeks.

### Approach 3b

`rd_approach3b_single` and `rd_approach3b_multiple` are the approach 3 columns
with the two `estrogen_progesterone` levels collapsed into one. A switch
between active levels never creates `previous`, so the relabel gives what a
full second pass would give.

```{r approach-3b}
episodes(skeleton[id == 16], "rd_approach3_single")
episodes(skeleton[id == 16], "rd_approach3b_single")
```

## 6. The exclusion flags

`ri_mht_excluded_product` is `TRUE` in every week of a person whose supplied
prescriptions include a flagged product, and `FALSE` otherwise.
`ri_mht_excluded_reason` carries the table's reason, or `NA`. Where one person
holds products flagged for different reasons, the reasons join in sorted
order, separated by `; `.

The flag is set from holding the product. It does not depend on that
prescription having a usable duration or a dispensing date, and it does not
depend on the week. A flagged prescription dispensed before or after
follow-up flags every observed week all the same. The prescriptions the
caller passes therefore define the window of ascertainment. The `ri_` prefix
means person level and time invariant.

The function removes nobody. It reports. A flagged person keeps every
exposure column, computed as for anyone else, and the caller decides what to
do with her rows.

```{r flags}
skeleton[, .N, keyby = .(ri_mht_excluded_product, ri_mht_excluded_reason)]
flagged <- skeleton[ri_mht_excluded_product == TRUE, unique(id)]
lmed[lopnr %in% flagged, .(lopnr, produkt, edatum, fddd)]
episodes(skeleton[id %in% flagged], "approach1")
```

Person 8 holds one Neofollin prescription among six others. Every one of her
312 weeks is flagged, and her approach columns are still written.

## 7. What the function does not do

- It removes no person. The exclusion flag is the caller's to act on.
- It reads no diagnosis and no covariate. The codebook's diagnosis-based
  exclusion is not applied. The table instead flags products whose holders the
  study removes.
- It builds no history before a woman's first observed week. A prescription
  dispensed before that week still covers the weeks of it inside follow-up,
  and the episode reads as starting in the first observed week.
- It imputes no duration, outside the six fixed-duration products. Any other
  prescription with no usable `fddd` contributes nothing.
- It reads the `Rules`, `FDDD` and `covariates` sheets of the codebook, and
  the `README` sheet of the product table, not at all.
